---
# --------------------------------------------
#
# IDAT_qc_report.Rmd
# Generate a QC report about IDAT files provided. 
#   This report is based on the minfi package and its qc functions.
# Version 1.0
# Date: 29 April 2022
# Alexis Hardy
# ULB 2022
#
# --------------------------------------------
#
# Steps:
# # Setup Rmd
# # Loading libraries, icons and input file
# # Generate QC data - detection p-value
# # Generate QC data - meth/unmeth signal
# # Generate QC data - control probes
# # Generate QC data - bval density
# # Print QC data and results - detection p-value
# # Print QC data and results - meth/unmeth signal
# # Print QC data and results - control probes
# # Print QC data and results - bval density
#
# --------------------------------------------

title: "IDAT processing - QC report"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
params:
  idat_dir: '../../../../GitHub_Files/GitHub/DNAm_Analysis/Idat_files/'
  control_path: '../../../../GitHub_Files/GitHub/DNAm_Analysis/DNAm_analysis_test/DATA/Control_population.csv.gz'
  truncate_samples_names: TRUE
  remove_probes_highpval: FALSE
  remove_samples_highpval: FALSE
  removeSamplesWithBadMethUnmeth: FALSE
  badMethUnmethSampleCutoff: 10.5
  remove_snip: FALSE
---

<!-- 2 HTML tags to define some styles that will be used in this Rmd -->
<style type="text/css">
h2 { /* Header 2 */
    font-size: 18px
}
</style>
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>

<!-- Setup Rmd -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`r paste0("Date of report: ", Sys.time())`

<!-- Loading libraries and input files (except annotation rds files). Input files are filtered according to provided parameters and prepared in a single format for all the Rmd. -->
<!-- All the following chunks are embedded into a HTML "details" flag to avoid warnings to be directly visible to the user but still accessible to check that no warning cause the report to be unreliable. -->
<details>
  <summary>Click here to show warnings</summary>
  <div class = "blue">
  
```{r loading_libraries_and_scripts, echo=FALSE, message=FALSE, collapse=TRUE}
###########################################################################
# Load required libraries
###########################################################################
library(minfi)#for QC pre-treatment
library(reshape2)#for QC pre-treatment
library(IlluminaHumanMethylationEPICmanifest)#@@@todo check for non-EPIC data
library(IlluminaHumanMethylation450kmanifest)#@@@todo check for non-450k data
library(kableExtra)
library(shiny)

###########################################################################
#Load scripts with functions
###########################################################################
source("func_qc.R")

###########################################################################
```
```{r loading_icons, echo=FALSE, message=FALSE, collapse=TRUE}
###########################################################################
# Generating shiny-icons
###########################################################################
iconGOOD <- icon("check-circle", style="color: rgb(100,190,0); font-size: 30px")
iconWARNING <- icon("question-circle", style="color: rgb(255,185,0); font-size: 30px")
iconBAD <- icon("times-circle", style="color: rgb(255,0,0); font-size: 30px")
iconDISABLED <- shiny::icon("minus-circle", style="color: rgb(200,200,200); font-size: 30px")

###########################################################################
```
```{r loading_data, echo=FALSE, message=FALSE, collapse=TRUE}
###########################################################################
# Loading idat files
###########################################################################
RGSet <- read.metharray.exp(params$idat_dir)
if(params$truncate_samples_names){
  sampleNames(RGSet) <- lapply(sampleNames(RGSet), {function(x) sub('_.*', '', x)})
}

###########################################################################
```
```{r check_detectionPVal, echo=FALSE, message=FALSE, collapse=TRUE}
###########################################################################
# Detection P-value computation and checking
###########################################################################

#Compute and format qc data - detection p-value -------------------------------------
data_detectpval <- get_qc_data_detectpval(RGSet)

#Detection P-value - get qc result -------------------------------------
DetectPvalQCresult <-
  get_qc_result_detectpval(data_detectpval$frac_failed_pos_per_sample,
                           data_detectpval$mean_pval_per_failed_pos,
                           data_detectpval$d_pval,
                           iconGOOD, iconWARNING, iconBAD,
                           params$remove_probes_highpval,
                           params$remove_samples_highpval)

###########################################################################
  
```
```{r check_MethUnmeth, echo=FALSE, message=FALSE, collapse=TRUE}

###########################################################################
# MethUnmeth median intensity comparison ----------------------------
###########################################################################

#Compute and format qc data - MethUnmeth median intensity -------------------------------------
data_methunmeth <- 
  get_qc_data_methunmeth(
    RGSet, 
    badSampleCutoff = as.numeric(params$badMethUnmethSampleCutoff)
    )

#MethUnmeth median intensity - get qc result -------------------------------------
MethUnmethQCresult <-
  get_qc_result_methunmeth(data_methunmeth$tblPlot_methUnmeth,
                           iconGOOD, iconBAD,
                           params$removeSamplesWithBadMethUnmeth)

###########################################################################

```
```{r check_ControlProbes, echo=FALSE, message=FALSE, collapse=TRUE}
###########################################################################
# Control Probes intensity ----------------------------
###########################################################################

# Compute and format qc data - Control probes  -------------------------------------
#   The following Code is an adapted version of qcReport function 
#   from minfi package.
data_ctrlprobes <- get_qc_data_ctrlprobes(RGSet)

#Control probes - get qc result -------------------------------------
ctrlProbesQCresult <- 
  get_qc_result_ctrlprobes(data_ctrlprobes$nb_bad_probes_per_sample, 
                           iconGOOD, iconWARNING)

###########################################################################
```
```{r check_BetaValueDensity, echo=FALSE, message=FALSE, collapse=TRUE}

###########################################################################
# Beta-value density ----------------------------
###########################################################################
bval_threshold = 250
plot_legend_text <- ""

if(!is.null(params$control_path)){
  # Compute and format qc data - Beta-value density  -------------------------------------
  data_bvaldensity <- get_qc_data_bvaldensity(RGSet, 
                                              params$control_path,
                                              bval_threshold)
  
  # Beta-value density - get qc result -------------------------------------
  bvalDensityDiffQCresult <- 
    get_qc_result_bvaldensity(
      data_bvaldensity$bval_abs_diff_per_sample,
      bval_threshold, 
      iconGOOD, iconWARNING
      )
  
  plot_legend_text <- 'Red curve is based on 
<span style=" color: white !important;border-radius: 4px; padding-right: 4px; padding-left: 4px; background-color: #FF0000 !important;" >the control population</span>
   provided.  '
  
} else {
  bvalDensityDiffQCresult <- list(
    iconDISABLED,
    paste0("Beta-value density Quality Check disabled.",
           " Please provide a control reference distribution",
           " to enable this part."),
    ""
    )
  
}

###########################################################################

```

  </div>
</details>


Icon | Description
:-------------:|-------------
`r iconGOOD`  | Good result. No need to perform an additional filter on this parameter.
`r iconWARNING`  | Some filter have been applied. User should check in this section if the sample(s) can be kept.
`r iconBAD` | Bad result. At least one sample do not meet the quality requirements and must be removed from the analysis.


***

# `r DetectPvalQCresult[[1]]` Detection p-value    
   
|   |
|:-------------:|
|  `r DetectPvalQCresult[[2]]`  |
|  `r DetectPvalQCresult[[3]]`  |
  
***  
  
  
##### Detection p-values distribution...  {.tabset .tabset-fade}  

  
###### for each sample
  
  
```{r detectPvalQCTable1AndPlot, echo=FALSE, fig.height=4, fig.width=6}
###########################################################################
# Barplot % failed positions per sample
###########################################################################

par(mar=c(5.1, 12.1, 0.1, 1.1))
frac_per_sample <- data_detectpval$frac_failed_pos_per_sample
barplot(frac_per_sample$FailedPositionsPct, 
        names.arg = frac_per_sample$sampleNames,
        col = ifelse(frac_per_sample$FailedPositionsPct > 1, 
                     "#FFA5A5", "grey"),
        xlim = c(0,max(frac_per_sample$FailedPositionsPct, 2)), 
        xlab = "Percentage of bad quality probes (%)", 
        horiz = TRUE, las=1)
abline(v=1, col="red", lty=2)
par(mar=c(5.1, 4.1, 4.1, 2.1))

###########################################################################
# Table % failed positions per sample
###########################################################################
data_detectpval$tbl_detectPVal1

###########################################################################
```
     


###### for each CpG position
  
```{r detectPvalQCTable2, echo=FALSE}
###########################################################################
# Table mean p-val per position for failed positions
###########################################################################
data_detectpval$tbl_detectPVal2

###########################################################################
```

#####

<br>
  
***

# `r MethUnmethQCresult[[1]]` Meth/Unmeth median intensity    
   
|   |
|:-------------:|
|  `r MethUnmethQCresult[[2]]`  |
|  `r MethUnmethQCresult[[3]]`  |
  
***  
  
  
```{r methUnmethQCTableAndPlot, echo=FALSE, fig.height=3.85, fig.width=5.5}
###########################################################################
# Plot Median Meth/Unmeth signal
###########################################################################
MethUnmethMedianPlot(data_methunmeth$tblPlot_methUnmeth,
                     params$badMethUnmethSampleCutoff)

###########################################################################
# Table Median Meth/Unmeth signal
###########################################################################
data_methunmeth$tbl_methUnmeth

###########################################################################
# Empty table to avoid overlapping due to "float" in the preceding table
###########################################################################
data_methunmeth$tblPlot_methUnmeth[1,, drop=FALSE] %>%
  kable("html") %>%
  kable_styling(full_width = TRUE) %>%
  row_spec(0:1, color = "transparent")

###########################################################################
```
     

<br>  

***  

# `r ctrlProbesQCresult[[1]]` Control probe signal intensity    
   
|   |
|:-------------:|
|  `r ctrlProbesQCresult[[2]]`  |
|  `r ctrlProbesQCresult[[3]]`  |
  
***  
  
  
```{r ctrlProbesQCTable, echo=FALSE, results = 'asis'}
###########################################################################
# Print table Number of “out-of-range” control probes per control type
###########################################################################
data_ctrlprobes$tbl_ctrlProbes1

###########################################################################
# Print table Number of “out-of-range” control probes per sample
###########################################################################
data_ctrlprobes$tbl_ctrlProbes2

###########################################################################
# Empty table to avoid overlapping due to "float" in the preceding table
###########################################################################
as.data.frame(data_ctrlprobes$nb_bad_probes_per_sample)[1,, drop=FALSE] %>%
  kable("html") %>%
  kable_styling(full_width = TRUE) %>%
  row_spec(0:1, color = "transparent")

###########################################################################
```
     
  
***  
  
##### Control probes...  {.tabset .tabset-fade}  

     
     
```{r ctrlProbesQCPlot, echo=FALSE, fig.height=5, fig.width=10, results = 'asis', message=FALSE}
###########################################################################
# Plot distribution of control probes signal per sample and per control type
###########################################################################

for (i in 1:length(data_ctrlprobes$controlsList)) {
  
  # Generate a tab
  cat(paste("######", names(data_ctrlprobes$controlsList)[i], "\n\n", sep=" "))
  flush.console()
  
  # Generate the plot
  controlProbesGreenRedPlot(data_ctrlprobes$controlsList[[i]], 
                            controlType=names(data_ctrlprobes$controlsList)[i])
  
  # Generate a space for a potential next tab
  cat("\n\n")
  flush.console()

  }

###########################################################################
```
     

#####

<br>  

***  

# `r bvalDensityDiffQCresult[[1]]` Beta-values distribution    
   
|   |
|:-------------:|
|  `r bvalDensityDiffQCresult[[2]]`  |
|  `r bvalDensityDiffQCresult[[3]]`  |
  
***  
    
`r plot_legend_text`    
  
   
  
```{r bvalDensityDiffQCTable, echo=FALSE, fig.height=3.85, fig.width=5.5, results = 'asis'}
###########################################################################
# If no control reference is provided, avoid plotting this part
###########################################################################

if(!is.null(params$control_path)){
  
  # Bval density plot
  par(mar=c(5.1, 4.1, 0.1, 2.1))
  bvalDensityPlot(data_bvaldensity$density_samples, 
                  data_bvaldensity$density_ref)
  par(mar=c(5.1, 4.1, 4.1, 2.1))
  
  # Print bval average difference with control reference
  print(data_bvaldensity$tbl_bvalDensityDiff)
  
  # Empty table to avoid overlapping due to "float" in the preceding table
  print(
  data_bvaldensity$bvalDensityDiff_table[1,, drop=FALSE] %>%
    kable("html") %>%
    kable_styling(full_width = TRUE) %>%
    row_spec(0:1, color = "transparent")
  )
}
###########################################################################

```
     

<br>  

*** 
  